### 2024-09-22 paper/paper.qmd
udpate

---
title: "Wasted Money: Toronto Police Budget and Crime Number"
subtitle: "Analyzing the changes in annual police budget and its effects
on the number of crimes within Toronto"
author: 
  - Yisu Hou
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(knitr)
library(kableExtra)

data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

```

# Introduction

You can and should cross-reference sections and sub-sections. We use @talia, @opendatatorotno, and @rohan.

The remainder of this paper is structured as follows. @sec-mydatasection

# Data {#sec-mydatasection}

Some of our data is of penguins (@fig-marriage), from @palmerpenguins.

```{r}
#| label: fig-marriage
#| fig-cap: Bills of penguins
#| echo: false

data |> 
  ggplot(aes(x = date, y = marriage_licenses)) +
  geom_point()

```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

\newpage

\appendix

# Appendix {-}

\newpage

# References
@r
@tidyverse
@dplyr
@knitr
@kextra
@opendatatoronto